MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver

### 2019-03-18 MAESTROeX_I/algorithm.tex
updates

There are a few key numerical modules we use in each time step.
\begin{itemize}
\item {\bf Average}$[\phi]\rightarrow[\overline\phi]$ computes the lateral average of a quantity over a layer at constant radius $r$, as described above in Section \ref{Sec:Spatial}.
\item {\bf Enforce HSE}$[\rho_0]\rightarrow[p_0]$ computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in Paper V).  The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
\MarginPar{add note about changes for irregular}
\item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc)]$ integrates the species and enthalpy due to reactions over $\Delta t/2$ by solving
\begin{equation}
\frac{dX_k}{dt} = \dot\omega_k(\rho,X_k,T); \qquad
\frac{dT}{dt} = \frac{1}{c_p}\left(-\sum_k\xi_k\dot\omega_k + H_{\rm nuc})\right)
\end{equation}
The inputs are the species, enthalpy, and base state pressure, and the outpus are the species, enthalpy, reaction rates, and nuclear energy generation rate.
See Papers III and V for details.
\end{itemize}

Each time step satisfies the standard advective CFL condition,
\begin{equation}
\Delta t = \sigma^{\rm CFL} \min_i(\Delta x / U_i),
\end{equation}
where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.5$.
There are additional constraints on the timestep that are typically much less restrictive than the advective CFL including
the acceleration due to the buoyancy force (sometimes in effect when the velocity is approximately zero at the start of some simulations) 
and the local magnitude of the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.

At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$...

\MarginPar{mention cutoff densities}

@@

\end{description}

This completes one step of the algorithm.\\ \\
{\em Initialization}
To initialize the simulation we use the same procedure described in Paper III.
We set initial values for $\Ub, \rho X_k$, and $\rho h$ and perform a sequence of projections 
(to ensure the velocity field satisfies the divergence constraint) 
followed by a small number of steps of the temporal advancement scheme to iteratively 
find initial values for $\pi^{n-\myhalf}$ and $S^0$ and $S^1$ for use in the first time step.

\subsection{Adaptive Mesh Refinement}
Talk about AMR algorithm and AMReX computational framework.

>>>>

There are a few key numerical modules we use in each time step.
\begin{itemize}
\item {\bf Average}$[\phi]\rightarrow[\overline\phi]$ computes the lateral average of a quantity over a layer at constant radius $r$, as described above in Section \ref{Sec:Spatial}.
\item {\bf Enforce HSE}$[\rho_0]\rightarrow[p_0]$ computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in Paper V).  The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
\MarginPar{add note about changes for irregular}
\item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc)]$ integrates the species and enthalpy due to reactions over $\Delta t/2$ by solving
\begin{equation}
\frac{dX_k}{dt} = \dot\omega_k(\rho,X_k,T); \qquad
\frac{dT}{dt} = \frac{1}{c_p}\left(-\sum_k\xi_k\dot\omega_k + H_{\rm nuc})\right)
\end{equation}
The inputs are the species, enthalpy, and base state pressure, and the outpus are the species, enthalpy, reaction rates, and nuclear energy generation rate.
See Papers III and V for details.
\end{itemize}
Each time step is constrained by the standard advective CFL condition,
\begin{equation}
\Delta t = \sigma^{\rm CFL} \min_i(\Delta x / U_i),
\end{equation}
where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.7$ and the minimum is taken over all spatial directions over all cells.
There are additional constraints on the timestep that are typically much less restrictive than the advective CFL including
the acceleration due to the buoyancy force (sometimes in effect when the velocity is approximately zero at the start of some simulations) 
and the local magnitude of the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.

To initialize the simulation we use the same procedure described in Paper III.
At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$.
We set initial values for $\Ub, \rho X_k$, and $\rho h$ and perform a sequence of projections 
(to ensure the velocity field satisfies the divergence constraint) 
followed by a small number of steps of the temporal advancement scheme to iteratively 
find initial values for $\pi^{n-\myhalf}$ and $S^0$ and $S^1$ for use in the first time step.
The new temporal integration strategy in MAESTROeX contains the following steps:
\MarginPar{mention cutoff densities}

@@

\end{description}
This completes one step of the algorithm.

Our approach to adaptive mesh refinement is algorithmically the same as the treatment described in Section 5 of Paper V; we refer the reader there for details.
We note that for spherical problems, AMR is only available for the case of a regularly-spaced base state.

### 2019-03-18 MAESTROeX_I/intro.tex
updates

We also present new algorithmic methodology that improves upon Paper V in a number of ways.
First, the overall temporal algorithm has been greatly simplified without compromising accuracy.
The key design decision was to replace the predicted evolution of the base state with 
a predictor-corrector approach.  Not only does this greatly simplify the dynamics of the base
state, but this treatment is more amenable to higher-order multiphysics coupling strategies
based on method-of-lines integration.  Second, we now provide additional spatial mapping
procedures that reduces, and in some cases eliminates, mapping error from the one-dimensional
base state to the Cartesian grid state for spherical problems.
Finally, the code has been completely ported to a new software framework suitable for modern manycore supercomputers.
The resulting MAESTROeX is implemented in the C++/F90 AMReX software library.  MAESTROeX uses MPI+OpenMP parallelism
and scales to XXX MPI processes, each capable of spawning tens of threads.
\MarginPar{quantify XXX.  I think it's 5,000-10,000 MPIs with potentially hundreds of threads}

>>>>

We also present new algorithmic methodology that improves upon Paper V in a number of ways.
First, the overall temporal algorithm has been greatly simplified without compromising accuracy.
The key design decision was to replace the predicted evolution of the base state with 
a predictor-corrector approach.  Not only does this greatly simplify the dynamics of the base
state, but this treatment is more amenable to higher-order multiphysics coupling strategies
based on method-of-lines integration.  Second, we now provide additional spatial mapping
procedures that reduces, and in some cases eliminates, mapping error from the one-dimensional
base state to the Cartesian grid state for spherical problems.
Finally, the code has been completely ported to a new software framework suitable for modern manycore supercomputers.
The resulting MAESTROeX is implemented in the C++/F90 AMReX software library.  MAESTROeX uses MPI+OpenMP parallelism
and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads
